---
title: "Lab-book"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "out") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyr)
library(lubridate)
library(summarytools)
library(survival)
library(survminer)
```

# About
This is your personal lab book and you may use it and structure it as you please. One way to structure it is to create a new section for each day, or maybe for each new function, or maybe both. It is up to you. Below is a fictitious example (that uses both), remove it when you have decided on a structure, you can find this code again in the `student-repo-template` repository on Github.

# Functions

## Load Data

Luckily our data is stored in a simple csv. If we are less lucky we might have to connect to SQL databases or work with weird formats. That can be done here.

```{r data, include=FALSE}
data <- read.csv("path_to_data.csv")
```

## Explore Data

Here we do some simple data exploration. This shows that... we will need to think about this when doing the analysis.

**TODO**:
- Use more complex methods for data exploration. Plot histograms for numeric features and bar plots for categorical? 

```{r explore, include=FALSE}
head(registry_data)
str(data)
summary(data)
stby(registry_data, by = NULL)
```

## Clean data

A few cases are missing data on `treatment_type` and `outcome_status`, we will remove these.

**TODO**:
- Exclude patients under the age of 18. I think dplyr can be used for that.

```{r explore, include=FALSE}
# Remove missing data
data_cleaned <- data %>%
  filter(!is.na(treatment_type) & !is.na(outcome_status))

print(paste("Number of rows removed during cleaning:", nrow(data) - nrow(data_cleaned)))
```

## Descriptive Statistics

Describe the distribution of patients among different treatment types.

```{r descriptive_stats, include=FALSE}
table(data_cleaned$treatment_type)
```

## Survival Analysis

```{r analysis, include=FALSE}
# Create a survival object
surv_obj <- Surv(data_cleaned$time_to_event, data_cleaned$event_occurred)

# Fit a survival curve based on treatment type
surv_fit <- survfit(surv_obj ~ treatment_type, data=data_cleaned)

# Plot the survival curve
ggsurvplot(surv_fit, data=data_cleaned,
           legend.title="Treatment Type",
           xlab="Days",
           ylab="Survival Probability",
           title="Survival Analysis by Treatment Type")
```

## Statistical Testing

We will use a log-rank test to compare survival curves for different treatment_types. We do this because of...

We will use a Cox Proportional Hazards regression model to understand how treatment type affects time to event while controlling for age and gender. The hazard ratios for each factor, especially treatment_type, are calculated to understand its impact on the survival time.

```{r analytical_stats, include=FALSE}
log_rank_test <- survdiff(surv_obj ~ treatment_type, data=data_cleaned)
print(log_rank_test)

cox_model <- coxph(formula = surv_obj ~ treatment_type + age + gender, data=data_cleaned)
summary(cox_model)
```

## Results and Visualisations

**TODO**:
- Create a Table one
- Create a table for hazard ratios
- Investigate if we can create a cool flowchart for inclusion/exclusion criteria

```{r results, include=FALSE}
data_cleaned %>%
  group_by(treatment_type) %>%
  tally() %>%
  ggplot(aes(x=treatment_type, y=n)) +
  geom_bar(stat="identity", fill="steelblue") +
  labs(title="Number of Patients by Treatment Type",
       x="Treatment Type",
       y="Number of Patients")

```

# Progress Report/Journal

## 10/8
Today i explored the possibility of using more advanced data visualisation techniques. The code below allows me to view numerical features using histograms and categorical using bar plots.

**TODO**:
- I should investigated if I can calculate hazard ratios for `treatment_type`, `age`, and `gender`.

```{r 10-8, include=FALSE}
# Exploring individual variables using histograms for numeric columns
numeric_columns <- sapply(registry_data, is.numeric)
for (col in names(registry_data)[numeric_columns]) {
  print(
    ggplot(registry_data, aes_string(col)) +
      geom_histogram(bins = 30, fill = "blue", alpha = 0.7) +
      theme_minimal() +
      labs(title = paste("Histogram of", col))
  )
}

# For categorical columns, using bar plots
factor_columns <- sapply(registry_data, is.factor)
for (col in names(registry_data)[factor_columns]) {
  print(
    ggplot(registry_data, aes_string(col)) +
      geom_bar(fill = "green", alpha = 0.7) +
      theme_minimal() +
      labs(title = paste("Bar plot of", col))
  )
}
```


## 14/8
Today I worked with calculating hazard ratios for `treatment_type`, `age`, and `gender`. After some research I figured out I need to use a Cox Proportional Hazards regression model to do this. I have a copied some code from the documentations below, it works with the current data.

**TODO**:
- Analyse the hazard ratios.

```{r 14-8, include=FALSE}
# Create a survival object
surv_obj <- Surv(data_cleaned$time_to_event, data_cleaned$event_occurred)

# Fit cox model to data
cox_model <- coxph(formula = surv_obj ~ treatment_type + age + gender, data=data_cleaned)

# Print summary of the regression model
summary(cox_model)
```

## 15/8

Today, I analysed the hazard ratios from the Cox Proportional Hazards regression model for `treatment_type`, `age`, and `gender`.

Findings:

- **Treatment Type**: HR of 1.5 (95% CI: [1.3, 1.7]). The new treatment increases the event risk by 50% compared to standard treatment.

- **Age**: HR of 1.02 (95% CI: [1.01, 1.03]). Each additional year of age increases the hazard rate by 2%.

- **Gender**: HR of 0.9 (95% CI: [0.8, 1.0]). Females have a 10% reduced hazard compared to males, but this result is borderline significant.

**TODO**:
- Visualize the hazard ratios using forest plots.
- Investigate predictor interactions.

```{r 15-8, include=FALSE}
# Extracting Hazard Ratios and their CI
hr_results <- tidy(cox_model, conf.int = TRUE, exponentiate = TRUE)

# Create a forest plot
ggforest(hr_results, data = data_cleaned)
```

